### 2017-04-23 paper/jfp-submission/safe.tex
more text

\paragraph{Dead Code}
%
Six programs (or 12\% the samples) contained type errors that were
unreachable by any inputs.
%
While technically safe, dead code is generally considered a ``code
smell''; students would likely benefit from a warning in this case.

\paragraph{Bad Function Call}
Fourteen programs contained a function call where the student had
supplied incorrect inputs, yet no error was triggered.
%
Consider, for example the following |assoc| function, which
looks up a key in an \emph{association list} and returns a default if
the key cannot be found.
%
\begin{code}
  let rec assoc (d, k, l) = match l with
    | (ki, vi)::tl ->
       if ki = k then
         vi
       else
         assoc (d, k, tl)
    | _ -> d

@@

>>>>

\paragraph{Dead Code}
%
Six programs (or 12\% the samples) contained type errors that were
unreachable by any inputs, often due to overlapping patterns in a
@match@ expression.
%
While technically safe, dead code is generally considered a ``code
smell''; students would likely benefit from a warning in this case.

\paragraph{Bad Function Calls}
Fourteen programs contained a function call where the student had
supplied incorrect inputs, yet no error was triggered.
%
Consider, for example the following |assoc| function, which
looks up a key in an \emph{association list} and returns a default if
the key cannot be found.
%
\begin{code}
  let rec assoc (d, k, l) = match l with
    | (ki, vi)::tl ->
       if ki = k then
         vi
       else
         assoc (d, k, tl)
    | _ -> d

@@

\paragraph{Non-Parametric Function Type}
%
Twenty-two programs (or 44\% of the samples) lack a witness in our
semantics due to our non-parametric $\tfun$ type for functions.
%
Recall that our goal is to expose the runtime errors that would have
been prevented by the type systems.
%
At runtime, it is always safe to call a function, thus we give functions
a simple type $\tfun$ that says they may be applied, but says nothing
about their inputs or outputs.
%
But consider the following @clone@ function, which is supposed to
produce a list containing @n@ copies of the input @x@.
%
% Seven programs violated the \emph{occurs check} with cyclic typing
% constraints like @'a = 'a list@, for example the following @clone@
% function.
%
\begin{code}
  let rec clone x n =
    if n > 0 then
      clone [x] (n - 1)
    else
      []
\end{code}
%
Unfortunately, the student instead constructs an @n@-level nested list
containing a single @x@.
%
The \ocaml compiler rejects this program because the recursive call to
@clone@ induces a cyclic typing constraint @'a = 'a list@, capturing the
fact that each call increases the nesting of the list.
%
\toolname fails to catch this because we do not track the types of the
inputs to @clone@.

We note, however, that @clone@ cannot go wrong; it is perfectly safe to
repeatedly enclose a list inside another (disregarding the fact that the
nested list is never returned).
%
Still, such a function would be very difficult to \emph{call} safely, as
the programmer would have to reason about the dependency between the
input @n@ and the nesting of the output list, which cannot be expressed
in \ocaml's type system.
%
Thus, is it not particularly satisfying that \toolname fails to produce
a witness here; a possible solution could be to track the types of the
inputs, and demonstrate to the user how they change between recursive
calls.
